\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{center}
	\Large{\textbf{ABSTRACT}}
\end{center}

The rapid development of implicit three-dimensional representations, such as Signed Distance Fields (SDFs) and neural fields, has significantly advanced automatic 3D reconstruction and generative modeling. Modern frameworks, including TRELLIS, are capable of producing high-quality implicit geometry that can be converted into polygonal meshes through differentiable isosurface extraction algorithms such as FlexiCubes. However, existing extraction methods primarily optimize geometric reconstruction accuracy and rarely consider the topological quality of the generated meshes. Consequently, the extracted meshes often exhibit irregular edge flow and connectivity, making them unsuitable for downstream applications such as 3D modeling, subdivision, rigging, and animation without an additional manual retopology process.

To address this limitation, this thesis proposes a topology-aware extension to the FlexiCubes isosurface extraction framework. The proposed approach first analyzes high-quality meshes from the ShapeNetCore dataset to derive local directional information that reflects desirable edge-flow characteristics. A Transformer-based neural network is then trained to predict surface normals and principal edge directions from sampled point clouds. During mesh extraction, the pretrained model is incorporated into the optimization process by introducing additional supervision on the predicted directions while jointly optimizing the signed distance field, vertex deformation, and FlexiCubes weighting parameters. This encourages the generated mesh to follow edge-flow patterns observed in artist-created models while preserving the geometric fidelity of the reconstructed surface.

Experimental evaluations were conducted on multiple 3D models generated by the TRELLIS framework. The proposed method was compared with the original FlexiCubes algorithm using geometric error metrics, including Root Mean Square (RMS) error, Hausdorff distance, and surface normal deviation under different mesh simplification levels. The results demonstrate that the proposed topology-aware optimization preserves reconstruction quality while producing meshes with more consistent edge-flow structures, reducing the need for manual retopology and improving the suitability of generated assets for practical content creation workflows.

\begin{flushright}
	\begin{tabular}{@{}c@{}}
		Student\\
		\textit{(Signature and full name)}
	\end{tabular}
\end{flushright}


\end{document}